---
title: "Lecture Example Interpolation - GGR376"
subtitle: Interpolation
output:
  pdf_document: default
  
---
```{r}
library(tidyverse)
library(leaflet)
library(rgdal)
```

# Interpolation

First, load some air pollution data from Paris France

```{r}

pollution <- readOGR("https://raw.githubusercontent.com/gisUTM/GGR376/master/Lab_2/paris_pollution.geojson")
```

Visualize the data on a leaflet map

```{r}
leaflet(data = pollution)%>% 
  addTiles()%>%
  addCircleMarkers()
              
```

There is an error. The Leaflet package expects all vector data to be specified in latitude and longitude using WGS 84 (a.k.a. EPSG:4326). By default, when displaying this data it projects everything to EPSG:3857 and expects that any map tiles are also displayed in EPSG:3857.

Check the pollution data's projection, which is stored in the sp object under proj4string

```{r}
pollution@proj4string
```

It is in a UTM projection.

Create a version of the data with the WGS 84 projection.

```{r}
pollution_LatLong <- spTransform(pollution, CRS("+proj=longlat +datum=WGS84 +no_defs"))

leaflet(pollution_LatLong)%>% 
  addTiles()%>%
  addCircleMarkers()
```

Keeping this in-mind, we should conduct our analysis in a projected coordinate system. Then we can project our final results for a final map.

## Theissen Polygons

Theissen polygons can be easily created with the voronoi function in the dismo library.

```{r}
library("dismo")

vor <- voronoi(pollution)
spplot(vor, "PM")

```

It is possible to add our points onto of our map. It is tricky, you need to create a list object with list(). Then provide the sp data, and any style you want to include.

```{r}
spplot(vor,"PM",
       sp.layout = list(
         pollution,
         col = "white",
         pch = 19)
       )
```

## Locations to predict

Read in this file of points that do not have pollution values. Use IDW to predict their values.

```{r}
predictions <- readOGR("https://raw.githubusercontent.com/gisUTM/GGR376/master/Lab_2/prediction_locations.geojson")
```

```{r}
plot(pollution)
plot(predictions, add = TRUE, col = "RED")
```


IDW Interpolation

The gstat library contains a set of functions for Spatial and Spatio-Temporal Geostatistical Modelling, Prediction and Simulation. 

```{r}
library(gstat)

pred <- idw(PM ~ 1, # Formula for IDW and Ordinary Kriging are Value ~ 1
            pollution, # Input Data
            predictions, # Prediction Locations
            idp = 2 # Exponent (k)
            )
```

The output is an object of class sp, with the prediction results stored in the variable: var1.pred

```{r}
class(pred)
str(pred)
spplot(pred, "var1.pred")
```

## Cross Validation

The gstat package has a built in cross validation function for both IDW and Kriging, krige.cv()  

The krige.cv function can conduct both LOOCV and k-fold validation

- LOOCV, set nfold = number of observations


```{r}
LOOCV <- krige.cv(PM~1, 
                  pollution, 
                  nfold = nrow(pollution), # Set the number of folds to the number of rows
                  set = list(idp = 2)
                  )
```

View the stucture of the object.

```{r}
str(LOOCV)
```

View the LOOCV data

```{r}
LOOCV@data
```


Plot the residuals.

```{r}
spplot(LOOCV, "residual")
```

Recall from lecture our Measures of Fit.

$$\operatorname{RMSE}=\sqrt{\frac{\sum_{i=1}^n (Y_i-\hat{Y_i})^2}{n}}$$

The $Y_i-\hat{Y_i}$ component is simply the values in the residual column . To calculate RMSE we will need to square our residuals, find the mean, and take the square.

```{r}
LOOCV@data$residual ^ 2 %>%
  mean() %>%
  sqrt()
  
```

This will need to be conducted a number of times. Here is a simple function to complete the same task.

```{r}
RMSE_resid <- function(x){
  return(sqrt(mean(x^2)))
} 
```

Try the function.

```{r}
RMSE_resid(LOOCV@data$residual)
```

At this point you could try different exponent values and compare the cross-validation to determine the optimal k.

```{r}
# Exponent of 1 (k)
LOOCV_k1 <- krige.cv(PM~1, 
                  pollution, 
                  nfold = nrow(pollution), # Set the number of folds to the number of rows
                  set = list(idp = 1)
                  )

# Exponent of 2 (k)
LOOCV_k2 <- krige.cv(PM~1, 
                  pollution, 
                  nfold = nrow(pollution), # Set the number of folds to the number of rows
                  set = list(idp = 2)
                  )

# Exponent of 3 (k)
LOOCV_k3 <- krige.cv(PM~1, 
                  pollution, 
                  nfold = nrow(pollution), # Set the number of folds to the number of rows
                  set = list(idp = 3)
                  )



```

Determine the best value of k, using RMSE of the residuals.
```{r}
# Determine RMSE when k = 1
RMSE_resid(LOOCV_k1@data$residual)

# Determine RMSE when k = 2
RMSE_resid(LOOCV_k2@data$residual)

# Determine RMSE when k = 3
RMSE_resid(LOOCV_k3@data$residual)
```



## Kriging

Kriging relies on a model to determine how the influence of points further away will occur in the estimation. Unlike IDW which has a constant rate of decay for distance.

Load some familiar libraries.

```{r}
library(sp)
library(gstat)
library(tidyverse)
```

The data set that will be used in this tutorial is meuse from the sp package.

"This data set gives locations and topsoil heavy metal concentrations, along with a number of soil and landscape variables at the observation locations, collected in a flood plain of the river Meuse, near the village of Stein (NL). Heavy metal concentrations are from composite samples of an area of approximately 15 m x 15 m."

```{r}
data(meuse)
glimpse(meuse) # Provide a nice view of our data
```

```{r}
class(meuse)
```

Take a look at the lead concentration in the soil. Create a plot using ggplot.

```{r}
ggplot(data = meuse)+
  geom_point(mapping = aes(x = x, y = y, size = lead), color = "brown", alpha = 3/4) +
  ggtitle("Topsoil Lead Concentrations (PPM)") + 
  coord_fixed()
```

The coord_fixed() function is added in ggplot to fix the scale along both the x and y axis. Remove this function and view the plot again. 

## Spatial Object

A spatial data object is required for gstat. Convert the data.frame into a SpatialPointsDataFrame().

This requires specifying the x and y coordinates using the sp::coordinates() function. A spatial object will be created from the data.frame

```{r}
coordinates(meuse) <- ~ x + y
class(meuse)
```

## Bounding Box

The spatial bounding box, which is a box that would surround all objects in your data, can be retrieved with sp::bbox()

```{r}
bbox(meuse)
```

## Fitting a Variogram

Kriging requires a variogram model, which is applied to determine the weighting of distance.

Kriging requires that our attribute be normally distributed.

```{r}
hist(meuse@data$lead)
```

As with most pollution data, the lead data is log-normal. When kriging it will be best to use the log of a value. Create a variable (logLead), which is the log values of lead.

```{r}
# Add logLead variable
meuse@data$logLead <- log(meuse@data$lead)

# Histogram of new variable
hist(meuse@data$logLead)
```

## Calculate Sample (empirical) Variogram

```{r}
lead.vgm <- variogram(logLead~1, meuse)
plot(lead.vgm)
```

## Variogram Models
There are a number of models available that can be used to model the data.

```{r}
show.vgms()
```

Select a model that has a similar shape to the data. In this case we will use "Sph", which is the Spherical Model.

```{r}
lead.fit <- fit.variogram(lead.vgm, model = vgm("Sph"))
```

The gstat::fit.variogram function fits ranges and/or sills from a simple or nested variogram model to a sample variogram. 

Recall

- Sill, value at which the model flattens out
- Range, distance at which model first flattens out
- Nugget, value at which the semi-variogram (almost) intercepts the y-value.
    + Expectation is that values at the same location should be equal
        + Due to sampling error or randomness they may not, which creates the nugget

## Plot the fit

```{r}
plot(lead.vgm, lead.fit)
```

The sill, nugget and range can be viewed by calling the fit model.

```{r}
lead.fit
```

- Sill: 0.515
- Range: 996.15
- Nugget: 0.051

## Create a convex hull

The boundary of the measured points represents the area that can be interpolated. Create a convex hull (polygon) around the points.

Find the row ids of the points that make up a convex hull
```{r}
convex_hull_point_ids <- meuse@coords %>% 
  chull() 

```

Subset the points that make up the convex hull, and change to a data frame
```{r}
convex_hull_points <- meuse@coords[ convex_hull_point_ids, ]%>%
  as.data.frame()

plot(convex_hull_points)
```

## Create a Polygon from the points

```{r}
# Get a matrix of the coordinates
crds <- coordinates(convex_hull_points)

# Create a polygon
poly <- sp::Polygon(crds)

# Prepare data for a SpatialPolygons Object
ID <- "Minimum Boundary"
Pls <- Polygons(list(poly), ID=ID)

SPls <- SpatialPolygons(list(Pls))
plot(SPls)
class(SPls)
```

## Create a evenly spaced grid

```{r}
grid <- makegrid(SPls, cellsize = 100)

class(grid)

str(grid)

coordinates(grid) <- ~ x1+x2

class(grid)

plot(grid)
```

## Select only grid points within Convex Hull

```{r}
# Check which grid cells are in the polygon
grid_in_poly <- sp::over(grid , SPls)

grid_in_poly # NA were not in poly

# Select points not NA using !is.na()
grid_predict <- grid[!is.na(grid_in_poly),]

plot(grid_predict)
```

## Predict Values at New Points

```{r}
lead_ordinary_krige <- krige(logLead~1, meuse, grid_predict, model=lead.fit)
```

## Plot the Predictions
```{r}
spplot(lead_ordinary_krige, "var1.pred")
```

## Plot the Kriging Variance

Estimate of estimation error. The error map is very much a map of distance to near points.

```{r}
spplot(lead_ordinary_krige, "var1.var")
```

```{r}
str(lead_ordinary_krige)
```

## Cross-Validation 

Use LOOCV

```{r}
LOOCV_krige_lead <-krige.cv(logLead~1, meuse, model=lead.fit )

LOOCV_krige_lead@data$residual
```

```{r}
# Note, this is in log units for lead
RMSE_resid(LOOCV_krige_lead@data$residual)
```









