---
title: "AndriyZvi - GGR376"
subtitle: Interpolation
output:
 
  
---

# Importing Libraries

```{r}
# Import libraries
library(tidyverse)
library(leaflet)
library(rgdal)
library(sp)
library(gapminder)
library(gstat)
library(spdep)
library(latticeExtra)
library(RColorBrewer)
library(dplyr)
library(rgeos)
library(sf)
```

# Load Data

```{r}
# Read in pollution data from csv
pollution<-read_csv("annual_conc_by_monitor_2018.csv")
# Load in state boundaries polygon
state_boundaries <- readOGR("state_boundaries.geojson")
```

# Data Filtering

```{r}
# Filter pollution data by state
pollution_filtered_states<-pollution%>%filter(`State Name`=="Arkansas"| `State Name`=="Tennessee" | `State Name`=="Louisiana" | `State Name`=="Oklahoma" | `State Name`== "Texas")

# Filter pollution data by sensor
pollution_filtered_states_by_parametres<-pollution_filtered_states%>%filter(`Parameter Name`=="Nitrogen dioxide (NO2)"| `Parameter Name`=="Ozone" | `Parameter Code`=="88101" & `Sample Duration`=="24-HR BLK AVG")

# Split up pollution data by sensor
no2 <- pollution_filtered_states%>%filter(`Parameter Name`=="Nitrogen dioxide (NO2)" & `Pollutant Standard`=="NO2 1-hour" & Datum=="WGS84")
o3 <- pollution_filtered_states%>%filter(`Parameter Name`=="Ozone" & `Pollutant Standard`=="Ozone 1-hour 1979" & Datum=="WGS84")
pm25 <- pollution_filtered_states%>%filter(`Parameter Code`=="88101" & `Sample Duration`=="24-HR BLK AVG" & `Pollutant Standard`=="PM25 24-hour 2012" & Datum=="WGS84")
no2_copy <- no2
o3_copy <- o3
pm25_copy <- pm25
```

#    Data Pre Plotting

```{r}
# All sensor data
ggplot(data = pollution_filtered_states_by_parametres)+
  geom_point(mapping = aes(x = Longitude, y = Latitude, size = `Arithmetic Mean`), color = "brown", alpha = 3/4) +
  ggtitle("All sensor Arithmetic Means") + 
  coord_fixed()

# NO2 sensors
ggplot(data = no2_copy)+
  geom_point(mapping = aes(x = Longitude, y = Latitude, size = `Arithmetic Mean`), color = "brown", alpha = 3/4) +
  ggtitle("NO2 sensor Arithmetic Means") + 
  coord_fixed()

# O3 sensors
ggplot(data = o3_copy)+
  geom_point(mapping = aes(x = Longitude, y = Latitude, size = `Arithmetic Mean`), color = "brown", alpha = 3/4) +
  ggtitle("O3 sensor Arithmetic Means") + 
  coord_fixed()

# PM 2.5 sensors
ggplot(data = pm25_copy)+
  geom_point(mapping = aes(x = Longitude, y = Latitude, size = `Arithmetic Mean`), color = "brown", alpha = 3/4) +
  ggtitle("PM 2.5 sensor Arithmetic Means") + 
  coord_fixed()
```

# Convert Data to SPDF and Project

```{r}
# NO2 data preview
data(no2)
glimpse(no2)
class(no2)
# NO2 convert to SPDF
coordinates(no2) <- ~ Longitude + Latitude

# O3 data preview
data(o3)
glimpse(o3)
class(o3)
# O3 convert to SPDF
coordinates(o3) <- ~ Longitude + Latitude

# PM 2.5 data preview
data(pm25)
glimpse(pm25)
class(pm25)
# PM 2.5 convert to SPDF
coordinates(pm25) <- ~ Longitude + Latitude

# Reproject data to WGS84 EPSG:4326
state_boundaries <- spTransform(state_boundaries, CRS("+init=epsg:4326"))
proj4string(no2) <- CRS("+init=epsg:4326")
no2 <- spTransform(no2, CRS("+init=epsg:4326"))
proj4string(o3) <- CRS("+init=epsg:4326")
o3 <- spTransform(o3, CRS("+init=epsg:4326"))
proj4string(pm25) <- CRS("+init=epsg:4326")
pm25 <- spTransform(pm25, CRS("+init=epsg:4326"))
```

#    Data Mapping

```{r}
# Map boundaries
leaflet()%>%
  addTiles()%>%
  addPolygons(data = state_boundaries, color = "#aaaaaa", weight = 1.0, smoothFactor = 0.5, opacity = 1.0, fillOpacity = 0.7)%>%
  addCircleMarkers(data = no2, color = "BLUE", weight = 1, radius = 2)%>%
  addCircleMarkers(data = o3, color = "GREEN", weight = 1, radius = 2)%>%
  addCircleMarkers(data = pm25, color = "RED", weight = 1, radius = 2)
```


#    Convex Hull pm25
```{r}
# convexhull_pm25 <- pm25@coords %>% 
#   chull() 
# convex_hull_pm25_points <- pm25@coords[ convexhull_pm25, ]%>%
#   as.data.frame()
# 
# plot(convex_hull_pm25_points)
# 
# crds <- coordinates(convex_hull_pm25_points)
# 
# poly <- sp::Polygon(crds)
# 
# ID <- "Minimum Boundary"
# Pls <- Polygons(list(poly), ID=ID)
```

#    Grid Making Convex Hull
```{r}
# SPls <- SpatialPolygons(list(SPls))
# plot(SPls)
# class(SPls)
```

```{r}
convexhull_o3 <- o3@coords %>% 
  chull() 
convex_hull_o3_points <- o3@coords[ convexhull_o3, ]%>%
  as.data.frame()

plot(convex_hull_o3_points)

crds <- coordinates(convex_hull_o3_points)

poly <- sp::Polygon(crds)

ID <- "Minimum Boundary"
Pls <- Polygons(list(poly), ID=ID)

SPls <- SpatialPolygons(list(Pls))
plot(SPls)
class(SPls)
```

```{r}
convexhull_no2 <- no2@coords %>% 
  chull() 
convex_hull_no2_points <- no2@coords[ convexhull_no2, ]%>%
  as.data.frame()

plot(convex_hull_no2_points)

crds <- coordinates(convex_hull_no2_points)

poly <- sp::Polygon(crds)

ID <- "Minimum Boundary"
Pls <- Polygons(list(poly), ID=ID)

SPls <- SpatialPolygons(list(Pls))
plot(SPls)
class(SPls)
```

```{r}
grid <- makegrid(SPls, cellsize = 100)
```

# Grid Making State Boundary

```{r}
# Plot sensors
# plot(no2, col = "BLUE")
# plot(o3, add = TRUE, col = "GREEN")
# plot(pm25, add = TRUE, col = "RED")

# Generate grid based on bounding box of state_boundaries
# state_boundaries <- unionSpatialPolygons(state_boundaries, NAME)
# state_boundaries <- aggregate(state_boundaries, dissolve = TRUE, FUN = sum)
# state_boundaries <- gUnaryUnion(state_boundaries, state_boundaries@data$NAME)


prediction_grid <- makegrid(state_boundaries, cellsize = 0.1)
class(prediction_grid)
str(prediction_grid)
coordinates(prediction_grid) <- ~ x1+x2
proj4string(prediction_grid) <- CRS("+init=epsg:4326")
class(prediction_grid)
# Add grid to plot
plot(prediction_grid)
plot(state_boundaries)
# Convert grid to SPDF
spatial_grid <- SpatialPoints(prediction_grid, proj4string = CRS(proj4string(state_boundaries)))
# Clip to boundary
spatial_grid_clipped <- prediction_grid[state_boundaries, ]

# # Check which grid cells are in the polygon
# overlap <- sp::over(prediction_grid , state_boundaries)
# overlap # NA were not in poly
# # Select points not NA using !is.na()
# spatial_grid_clipped <- prediction_grid[!is.na(overlap),]
# plot(spatial_grid_clipped)

# Plot spatial grid with sensors
# plot(spatial_grid_clipped)
# plot(no2, add = TRUE, col = "BLUE")
# plot(o3, add = TRUE, col = "GREEN")
# plot(pm25, add = TRUE, col = "RED")
```

#IDW OF PM2.5
```{r}
idw_pm25 <- idw(pm25@data$`Arithmetic Mean` ~ 1, # Formula for IDW and Ordinary Kriging are Value ~ 1
            pm25, # Input Data
            spatial_grid_clipped, # Prediction Locations
            idp = 2 # Exponent (k)
            )

class(idw_pm25)
str(idw_pm25)
spplot(idw_pm25, "var1.pred")
```
#PM25 LOOCV K=2 (Too High)
```{r}
pm25@data$AM <- pm25@data$'Arithmetic Mean'
LOOCV_PM25_IDW <- krige.cv(AM ~ 1,
                  pm25,
                  nfold = nrow(pm25),
                  set = list(idp = 2)
                  )

str(LOOCV_PM25_IDW)
LOOCV_PM25_IDW@data
spplot(LOOCV_PM25_IDW, "residual")

LOOCV_PM25_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()


```
#PM25 LOOCV K=3 (Just Right)
```{r}
pm25@data$AM <- pm25@data$'Arithmetic Mean'
LOOCV_PM25_IDW<- krige.cv(AM ~ 1,
                  pm25,
                  nfold = nrow(pm25),
                  set = list(idp = 3)
                  )

str(LOOCV_PM25_IDW)
LOOCV_PM25_IDW@data
spplot(LOOCV_PM25_IDW, "residual")

LOOCV_PM25_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()

```
##PM25 LOOCV 2.5 K4(Too high)
```{r}
pm25@data$AM <- pm25@data$'Arithmetic Mean'
LOOCV_PM25_IDW <- krige.cv(AM ~ 1,
                  pm25,
                  nfold = nrow(pm25),
                  set = list(idp = 4)
                  )

str(LOOCV_PM25_IDW)
LOOCV_PM25_IDW@data
spplot(LOOCV_PM25_IDW, "residual")

LOOCV_PM25_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()

RMSE_resid <- function(x){
  return(sqrt(mean(x^2)))
} 
RMSE_resid(LOOCV@data$residual)
```


#IDW Interpolation Ozone
```{r}
idw_o3 <- idw(o3@data$`Arithmetic Mean` ~ 1, # Formula for IDW and Ordinary Kriging are Value ~ 1
            o3, # Input Data
            spatial_grid_clipped, # Prediction Locations
            idp = 2 # Exponent (k)
            )

class(idw_o3)
str(idw_o3)
spplot(idw_o3, "var1.pred")
```
#Ozone LOOCV 2
```{r}
o3@data$AM <- o3@data$'Arithmetic Mean'
LOOCV_o3_IDW <- krige.cv(AM ~ 1,
                  o3,
                  nfold = nrow(o3),
                  set = list(idp = 2)
                  )

str(LOOCV_o3_IDW)
LOOCV_o3_IDW@data
spplot(LOOCV_o3_IDW, "residual")

LOOCV_o3_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()

```
#IDW Ozone LOOCV 3(just right)
```{r}
o3@data$AM <- o3@data$'Arithmetic Mean'
LOOCV_o3_IDW <- krige.cv(AM ~ 1,
                  o3,
                  nfold = nrow(o3),
                  set = list(idp = 3)
                  )

str(LOOCV_o3_IDW)
LOOCV_o3_IDW@data
spplot(LOOCV_o3_IDW, "residual")

LOOCV_o3_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()

```
##IDW LOOCV Ozone 4
```{r}
o3@data$AM <- o3@data$'Arithmetic Mean'
LOOCV_o3_IDW <- krige.cv(AM ~ 1,
                  o3,
                  nfold = nrow(o3),
                  set = list(idp = 4)
                  )

str(LOOCV_o3_IDW)
LOOCV_o3_IDW@data
spplot(LOOCV_o3_IDW, "residual")

LOOCV_o3_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()
```


##IDW OF NITROGEN OXIDE
```{r}
idw_no2 <- idw(no2@data$`Arithmetic Mean` ~ 1, # Formula for IDW and Ordinary Kriging are Value ~ 1
            no2, # Input Data
            spatial_grid_clipped, # Prediction Locations
            idp = 2 # Exponent (k)
            )

class(idw_no2)
str(idw_no2)
spplot(idw_no2, "var1.pred")
```

#LOOCV IDW Nitrogen Oxide 2
```{r}
no2@data$AM <- no2@data$'Arithmetic Mean'
LOOCV_no2_IDW <- krige.cv(AM ~ 1,
                  no2,
                  nfold = nrow(no2),
                  set = list(idp = 2)
                  )

# TEST LOOCV CODE
# # Preview LOOCV
# str(LOOCV)
# LOOCV@data
# # Plot LOOCV
# spplot(LOOCV, "residual")
# 
# # Measure of Fit
# LOOCV@data$residual ^ 2 %>%
#   mean() %>%
#   sqrt()
# # Define function to calculate measure of fit for all
# RMSE_resid <- function(x){
#   return(sqrt(mean(x^2)))
# }
# # Run function
# RMSE_resid(LOOCV@data$residual)

str(LOOCV_no2_IDW)
LOOCV_no2_IDW@data
spplot(LOOCV_no2_IDW, "residual")

LOOCV_no2_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()
```
## LOOCV 3 Nitrogen Oxide (Higher)
```{r}
no2@data$AM <- no2@data$'Arithmetic Mean'
LOOCV_no2_IDW <- krige.cv(AM ~ 1,
                  no2,
                  nfold = nrow(no2),
                  set = list(idp = 3)
                  )

str(LOOCV_no2_IDW)
LOOCV_no2_IDW@data
spplot(LOOCV_no2_IDW, "residual")

LOOCV_no2_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()
```
## LOOCV Nitrogen Oxide 1 (LOWER)
```{r}
no2@data$AM <- no2@data$'Arithmetic Mean'
LOOCV_no2_IDW <- krige.cv(AM ~ 1,
                  no2,
                  nfold = nrow(no2),
                  set = list(idp = 1)
                  )

str(LOOCV_no2_IDW)
LOOCV_no2_IDW@data
spplot(LOOCV_no2_IDW, "residual")

LOOCV_no2_IDW@data$residual ^ 2 %>%
  mean() %>%
  sqrt()
```

# Kriging
# Data Transforming

```{r}
# NO2 pre transform
class(no2)
bbox(no2)
hist(no2@data$'Arithmetic Mean')
# Log transform NO2 data to meet normally distributed assumption for kriging
no2@data$AMLog <- log(no2@data$'Arithmetic Mean')
# Histogram of new variable
hist(no2@data$AMLog)
# Show variogram models
show.vgms()
# NO2 variogram
no2.vgm <- variogram(AMLog~1, no2)
plot(no2.vgm)

# O3 pre transform
class(o3)
bbox(o3)
hist(o3@data$'Arithmetic Mean')
# Log transform O3 data to meet normally distributed assumption for kriging
o3@data$'AMLog' <- log(o3@data$'Arithmetic Mean')
# Histogram of new variable
hist(o3@data$AMLog)
# Show variogram models
show.vgms()
# O3 variogram
o3.vgm <- variogram(AMLog~1, o3)
plot(o3.vgm)

# PM 2.5 pre transform
class(pm25)
bbox(pm25)
hist(pm25@data$'Arithmetic Mean')
# Log transform PM 2.5 data to meet normally distributed assumption for kriging
pm25@data$AMLog <- pm25@data$'Arithmetic Mean'
# Histogram of new variable
hist(pm25@data$AMLog)
# Show variogram models
show.vgms()
# PM 2.5 variogram
pm25.vgm <- variogram(AMLog~1, pm25)
plot(pm25.vgm)
```

# Fit Variogram

```{r}
# Fit NO2 data
no2.fit <- fit.variogram(no2.vgm, model = vgm("Pen"))
plot(no2.vgm, no2.fit)
# Fit O3 data
o3.fit <- fit.variogram(o3.vgm, model = vgm("Bes"))
plot(o3.vgm, o3.fit)
# Fit PM 2.5 data
pm25.fit <- fit.variogram(pm25.vgm, model = vgm("Gau"))
plot(pm25.vgm, pm25.fit)
```

# Predict Kriging Values

```{r}
no2_ordinary_krige <- krige(AMLog~1, no2, spatial_grid_clipped, model=no2.fit)
o3_ordinary_krige <- krige(AMLog~1, o3, spatial_grid_clipped, model=o3.fit)
pm25 = pm25[which(!duplicated(pm25@coords)), ]
pm25_ordinary_krige <- krige(AMLog~1, pm25, spatial_grid_clipped, model=pm25.fit)

# # pm25_fixed <- pm25[-zerodist(pm25)[,1],]
# # zerodist(pm25_fixed)
# 
# # pm25_fixed <- pm25[which(!duplicated(pm25@data$`Site Num`)), ]
# 
# pm25_fixed = pm25[which(!duplicated(pm25@coords)), ]
# 
# #pm25_ordinary_krige <- krige.cv(AMLog~1,pm25)
# pm25_ordinary_krige <- krige.cv(AMLog~1, pm25_fixed, spatial_grid_clipped, model=pm25.fit)
# #pm25_ordinary_krige <- krige.cv(AMLog~1, pm25, spatial_grid_clipped, model=pm25.fit)
```

```{r}
spplot(no2_ordinary_krige, "var1.pred")
str(no2_ordinary_krige)
LOOCV_krige_no2 <-krige.cv(AMLog~1, no2, model=no2.fit)
LOOCV_krige_no2@data$residual
RMSE_resid(LOOCV_krige_no2@data$residual)

spplot(o3_ordinary_krige, "var1.pred")
str(o3_ordinary_krige)
LOOCV_krige_o3 <-krige.cv(AMLog~1, o3, model=o3.fit)
LOOCV_krige_o3@data$residual
RMSE_resid(LOOCV_krige_o3@data$residual)

spplot(pm25_ordinary_krige, "var1.pred")
str(pm25_ordinary_krige)
LOOCV_krige_pm25 <-krige.cv(AMLog~1, pm25, model=pm25.fit)
LOOCV_krige_pm25@data$residual
RMSE_resid(LOOCV_krige_pm25@data$residual)
```

# Spatial Clustering

```{r}
# p2g <- points2grid(points = spatial_grid_clipped)
# g2n <- grid2nb(grid = p2g, d = p2g@cells.dim)
# lcosts_idw <- nbcosts(g2n, dpad_idw)
# state_boundaries <- st_read("state_boundaries.shp")

class(no2_ordinary_krige)

dpad_krige <- cbind(idw_no2@data$var1.pred, idw_o3@data$var1.pred, idw_pm25@data$var1.pred)
#dpad_krige <- data.frame(scale(as.data.frame(state_boundaries)[,5:8]))
summary(dpad_krige)

krige_nb <- poly2nb(state_boundaries)
lcosts_krige <- nbcosts(krige_nb, dpad_krige)
krige_w <- nb2listw(krige_nb, lcosts_krige, style="B")
mst_krige <- mstree(krige_w,5)

par(mar=c(0,0,0,0))
plot(state_boundaries, border=gray(.5))
plot(mst_krige, coordinates(state_boundaries), col=2, 
     cex.lab=.6, cex.circles=0.035, fg="blue", add=TRUE)

### three groups with no restriction
res1 <- skater(mst_krige[,1:2], dpad_krige, 2)

### groups size
table(res1$groups)

### the skater plot
opar <- par(mar=c(0,0,0,0))
plot(res1, coordinates(state_boundaries), cex.circles=0.035, cex.lab=.7)

### the skater plot, using other colors
plot(res1, coordinates(state_boundaries), cex.circles=0.035, cex.lab=.7,
     groups.colors=heat.colors(length(res1$ed)))

### the Spatial Polygons plot
plot(state_boundaries, col=heat.colors(length(res1$edg))[res1$groups])
```

